---
title: "Project File"
output: html_notebook
---


```{r}
library(forecast)
library(zoo)
library(fpp3)
library(tsibbledata)
library(tsibble)

rm(list=ls())
```


Units Sold Data 
```{r}
sold1 <- read.csv('Units Sold 1.csv')
sold2 <- read.csv('Units Sold 2.csv')
#missing Units Sold 3 documentation 
sold4 <- read.csv('Units Sold 4.csv')

sold1
```
Claims Data 
```{r}
claims1 <- read.csv('Claims 1.csv')
claims2 <- read.csv('Claims 2.csv')
claims3 <- read.csv('Claims 3.csv')
claims4 <- read.csv('Claims 4.csv')
```

Ratings Data 
```{r}
ratings1 <- read.csv('Ratings 1.csv')
ratings2 <- read.csv('Ratings 2.csv')
ratings3 <- read.csv('Ratings 3.csv')
ratings1
```


Convert to ts object 
Q: WHAT YEAR DO THESE START ACCUMULATING DATA? THIS DATA IS SINCE THEY STARTING SELLING BUT WHEN WAS THAT FOR EACH PRODUCT? 
```{r}
sold1.ts <- ts(sold1$Sum.of.TotalUnitsSold, freq = 1)
sold2.ts <- ts(sold2$Sum.of.TotalUnitsSold)
sold4.ts <- ts(sold4$Sum.of.TotalUnitsSold)

sold1.ts
```

```{r}
claims1.ts <- ts(claims1$Total.Claims)
claims2.ts <- ts(claims2$Total.Claims)
claims3.ts <- ts(claims3$Total.Claims)
claims4.ts <- ts(claims4$Total.Claims)
```


Plot time series for all products units sold and claims to get an idea of trends, seasonality, etc. 

```{r}
plot(sold1.ts, ylim = c(1000,150000), ylab = "Units Sold", xlab = "Time (months from launch)", main = 'Units Sold Over Time')
lines(sold2.ts, col = 'green')
lines(sold4.ts, col = 'blue')

```


```{r}
plot(claims1.ts, ylim = c(0, 3000), ylab = "Number of Claims Received", xlab = 'Time (months from launch)', main = 'Claims Received Over Time')
lines(claims2.ts, col = 'green')
lines(claims3.ts, col = 'red')
lines(claims4.ts, col = 'blue')

```

1. Deep Dive- Product 1

a) Split Train/Test
```{r}
#split train/test - using the last 10 months as validation (we can make this however large)
validLength = 10
nTrain <- length(sold1.ts) - validLength
train.ts <- window(sold1.ts, start = c(1, 1), end = c(1, nTrain))
valid.ts <- window(sold1.ts, start = c(1, nTrain + 1), end = c(1, nTrain + validLength))
```

b) Plotting ACF of ts data 
```{r}
Acf(sold1.ts)
```

c) Smoothing Methods 
```{r}
plot(sold1.ts, ylim = c(1000,150000), ylab = "Units Sold", xlab = "Time (months from launch)", main = 'Units Sold Over Time', xaxt = 'n')
axis(1, at = seq(1, 45,1), labels = format(seq(1,45,1)))
lines(c(nTrain,nTrain), c(0,150000))
text(nTrain-3, 140000, "Train")
text(nTrain+3, 140000, 'Test')

```
Try to use Moving Average 
```{r}
ma.trailing <- rollmean(train.ts, k = 12, align = 'right')
last.ma <- tail(ma.trailing,1)
ma.trailing.pred <- ts(rep(last.ma, validLength), start=c(1,nTrain+1), end = c(1, nTrain+validLength),freq=1)
```

```{r}
plot(sold1.ts, ylim = c(1000,150000), ylab = "Units Sold", xlab = "Time (months from launch)", main = 'Units Sold Over Time', xaxt = 'n')
axis(1, at = seq(1, 45,1), labels = format(seq(1,45,1)))
lines(ma.trailing, col='blue')
lines(ma.trailing.pred, col = 'blue', lty =2)
lines(c(nTrain,nTrain), c(0,150000))
text(nTrain-3, 140000, "Train")
text(nTrain+3, 140000, 'Test')

```

```{r}
accuracy(ma.trailing.pred, valid.ts)

```


Try to use Simple Exponential Smoothing 
```{r}
ses <- ets(train.ts, model = 'ANN')
ses
ses.pred <- forecast(ses, h = validLength, level =0)
ses.pred
```

```{r}
plot(sold1.ts, ylim = c(1000,150000), ylab = "Units Sold", xlab = "Time (months from launch)", main = 'Units Sold Over Time', xaxt = 'n')
axis(1, at = seq(1, 45,1), labels = format(seq(1,45,1)))
lines(ses.pred$mean, lwd = 2, col = 'blue', lty =2) 
lines(c(nTrain,nTrain), c(0,150000))
text(nTrain-3, 140000, "Train")
text(nTrain+3, 140000, 'Test')
```

```{r}
accuracy(ses.pred$mean, valid.ts)
```

Try to use Double Exponential Smoothing 
```{r}
des <- ets(y = train.ts, model = "ZAN")
des.pred <- forecast(des, h = validLength, level = 0)
des.pred
```

```{r}
plot(sold1.ts, ylim = c(1000,150000), ylab = "Units Sold", xlab = "Time (months from launch)", main = 'Units Sold Over Time', xaxt = 'n')
axis(1, at = seq(1, 45,1), labels = format(seq(1,45,1)))
lines(des.pred$mean, lwd = 2, col = 'blue', lty =2) 
lines(c(nTrain,nTrain), c(0,150000))
text(nTrain-3, 140000, "Train")
text(nTrain+3, 140000, 'Test')

```

```{r}
accuracy(des.pred$mean, valid.ts)

```


Letting R choose ETS model 
```{r}
chosen_model <- ets(train.ts)
chosen_model

#model chosen is double exponential smoothing: ETS(M,A,N)
```

Regression Methods 

Trend Model
```{r}
train.trend.lm <- tslm(train.ts ~trend)

plot(train.ts, xlab = 'Time', ylab = 'Ridership')
lines(train.trend.lm$fitted, lwd=2, col = 'blue')


```

```{r}
summary(train.trend.lm)

```
```{r}
train.trend.lm.pred <- forecast(train.trend.lm, h = validLength, level =0)

plot(sold1.ts, ylim = c(1000,150000), ylab = "Units Sold", xlab = "Time (months from launch)", main = 'Units Sold Over Time', xaxt = 'n')
axis(1, at = seq(1, 45,1), labels = format(seq(1,45,1)))
lines(train.trend.lm$fitted, lwd=2, col = 'blue')
lines(train.trend.lm.pred$mean, lty = 2, col = 'blue')
lines(c(nTrain,nTrain), c(0,150000))
text(nTrain-3, 140000, "Train")
text(nTrain+3, 140000, 'Test')

```


Quadratic Trend Model 
```{r}
train.quad.lm <- tslm(train.ts ~ trend + I(trend^2))

plot(train.ts, xlab = 'Time', ylab = 'Ridership')
lines(train.quad.lm$fitted, lwd=2, col = 'blue')
```

```{r}
summary(train.quad.lm)

```


```{r}
train.quad.lm.pred <- forecast(train.quad.lm, h = validLength, level =0)

plot(sold1.ts, ylim = c(1000,150000), ylab = "Units Sold", xlab = "Time (months from launch)", main = 'Units Sold Over Time', xaxt = 'n')
axis(1, at = seq(1, 45,1), labels = format(seq(1,45,1)))
lines(train.trend.lm$fitted, lwd=2, col = 'blue')
lines(train.quad.lm.pred$mean, lty = 2, col = 'blue')
lines(c(nTrain,nTrain), c(0,150000))
text(nTrain-3, 140000, "Train")
text(nTrain+3, 140000, 'Test')


```
Exponential Trend Model 
```{r}
train.exp.lm <- tslm(train.ts ~ trend, lambda =0)

plot(train.ts, xlab = 'Time', ylab = 'Ridership')
lines(train.exp.lm$fitted, lwd=2, col = 'blue')

```

```{r}
summary(train.exp.lm)

```

```{r}
train.exp.lm.pred <- forecast(train.exp.lm, h = validLength, level =0)

plot(sold1.ts, ylim = c(1000,150000), ylab = "Units Sold", xlab = "Time (months from launch)", main = 'Units Sold Over Time', xaxt = 'n')
axis(1, at = seq(1, 45,1), labels = format(seq(1,45,1)))
lines(train.trend.lm$fitted, lwd=2, col = 'blue')
lines(train.exp.lm.pred$mean, lty = 2, col = 'blue')
lines(c(nTrain,nTrain), c(0,150000))
text(nTrain-3, 140000, "Train")
text(nTrain+3, 140000, 'Test')


```

Additive Seasonal Model with Trend 
```{r}
train.add.trend.lm <- tslm(train.ts ~trend + season)

plot(train.ts, xlab = 'Time', ylab = 'Ridership')
lines(train.add.trend.lm$fitted, lwd=2, col = 'blue')

```


